---
title: "Checking completeness of data"
output: html_notebook
---

## Preliminary

```{r}
suppressPackageStartupMessages(library(tidyverse))
library(formattable)
```

Read in the relevant data

```{r, message=FALSE}
meae_maps <- read_csv("elections-data/maps.csv")
meae_maps_to_elections <- read_csv("elections-data/maps-to-elections.csv")
meae_elections <- read_csv("elections-data/elections.csv")
meae_congress <- read_csv("elections-data/congressional-counties.csv")
```

## Checking maps and maps-to-elections

Jordan has completed the list of maps to elections. How many maps are there? These two should be equal.

```{r}
meae_maps %>% nrow()
meae_maps_to_elections %>% count(meae_id) %>% nrow()
```

Also, an anti-join of these two tables should return 0 mismatches when run in both directions.

```{r}
meae_maps %>% 
  anti_join(meae_maps_to_elections, by = "meae_id")

meae_maps_to_elections %>% 
  anti_join(meae_maps, by = "meae_id")
```

## Checking completeness of maps by state and Congress

Now we want to check that there contiguous runs for each state. In other words, once a state has a congression election, it should keep having the congressional elections.

```{r}
meae_maps %>% 
  select(state, congress) %>% 
  mutate(available = "✔") %>% 
  spread(state, available, fill = "") %>% 
  knitr::kable()
```

## Checking maps-to-elections and elections

Now how many elections IDs that are present in the maps-to-elections table are not present in the elections table?

```{r}
meae_maps %>% 
  left_join(meae_maps_to_elections, by = "meae_id") %>% 
  anti_join(meae_elections, by = "election_id") %>% 
  select(election_id, meae_id, everything())
```

Previously there was a problem here but Jordan fixed it.

## Checking how many maps have data available

We will say that a map has data available if there is at least one non-zero vote recorded for that election (in the NNV sense).

```{r}
elections_present <- meae_congress %>% 
  filter(vote > 0) %>% 
  count(election_id)

meae_maps %>% 
  left_join(meae_maps_to_elections, by = "meae_id") %>% 
  left_join(meae_elections, by = "election_id") %>% 
  anti_join(elections_present, by = "election_id") %>% 
  rename(state = state.x, congress = congress.x) %>% 
  count(meae_id, state, congress) %>% 
  arrange(desc(n)) %>% 
  select(-meae_id) %>% 
  spread(state, n, fill = "") %>% 
  knitr::kable()
```

Now let's get a list of those missing NNV election IDs.

```{r}
elections_missing_data <- meae_maps %>% 
  left_join(meae_maps_to_elections, by = "meae_id") %>% 
  left_join(meae_elections, by = "election_id") %>% 
  anti_join(elections_present, by = "election_id") %>% 
  rename(state = state.x, congress = congress.x) %>% 
  select(-state.y, -congress.y) %>% 
  arrange(state, congress)

elections_missing_data
write_csv(elections_missing_data,
          "~/dev/mapping-elections/data-cleaning/missing-congressional-counties/elections-missing-data.csv")
```


