---
title: "Checking completeness of data"
output: html_notebook
---

```{r}
suppressPackageStartupMessages(library(tidyverse))
library(formattable)
```

Read in the relevant data

```{r}
meae_maps <- read_csv("elections-data/maps.csv")
meae_maps_to_elections <- read_csv("elections-data/maps-to-elections.csv")
meae_elections <- read_csv("elections-data/elections.csv")
```


Jordan has completed the list of maps to elections. How many maps are there? These two should be equal.

```{r}
meae_maps %>% nrow()
meae_maps_to_elections %>% count(meae_id) %>% nrow()
```

Also, an anti-join of these two tables should return 0 mismatches when run in both directions.

```{r}
meae_maps %>% 
  anti_join(meae_maps_to_elections, by = "meae_id")

meae_maps_to_elections %>% 
  anti_join(meae_maps, by = "meae_id")
```

Now we want to check that there contiguous runs for each state. In other words, once a state has a congression election, it should keep having the congressional elections.

```{r}
meae_maps %>% 
  select(state, congress) %>% 
  mutate(available = "✔") %>% 
  spread(state, available, fill = "") %>% 
  knitr::kable()
```

Now how many elections IDs that are present in the maps-to-elections table are not present in the elections table?

```{r}
meae_maps %>% 
  left_join(meae_maps_to_elections, by = "meae_id") %>% 
  anti_join(meae_elections, by = "election_id") %>% 
  select(election_id, meae_id, everything())
```

